---
title: "Hans Rosling's Human Development Plot (Daniel Posmik)"
author: "Daniel Posmik"
date: "10/16/2021"
output:
  html_document: default
  pdf_document: default
---

## R Setup

```{r, message = FALSE}
#install.packages("gapminder")
library(gapminder)
library(ggplot2)
```

First, let us look at the visualization for the years 1952 - 2007.
``` {r, message = FALSE}
#install.packages("plotly")
library(plotly) #We load and install the plotly package

g <- crosstalk::SharedData$new(gapminder, ~continent) #Creates slider widget

gg <- ggplot(g, aes(gdpPercap, lifeExp, color = continent, frame = year)) +
  #gdpPercap is on the x-axis and lifeExp is on the y-axis; color represents continent
  geom_point(aes(size = pop, ids = country)) + 
  #size of bubbles represents population by country
  #geom_smooth(se = FALSE, method = "lm") + #this is an optional smoothing line
  scale_x_log10() #this represents the log scale

ggplotly(gg) %>% 
  highlight("plotly_hover")
```
We can see that the plot is very similar to the one on the website. However, now we need to rename the axes.

Since I have already visualized "at least three years" (per the instructions), I decide against further limiting the time frame of this visualization for simplicity. 

``` {r, message = FALSE}
#install.packages("scales")
library(scales)

g <- crosstalk::SharedData$new(gapminder, ~continent) #Creates slider widget

gg <- ggplot(g, aes(gdpPercap, lifeExp, color = continent, frame = year)) +
  #gdpPercap is on the x-axis and lifeExp is on the y-axis; color represents continent
  geom_point(aes(size = pop, ids = country)) + 
  #size of bubbles represents population by country
  #geom_smooth(se = FALSE, method = "lm") + #this is an optional smoothing line
  scale_x_log10() + #this represents the log scale
  labs(x="Income") + #Rename x-axis using scales package
  labs(y="Life Expectancy") #Rename y-axis using scales package

ggplotly(gg) %>% 
  highlight("plotly_hover")
```
Finally, we adjust the visuals to fit the example as closely as possible. 

``` {r, message = FALSE}
#install.packages("scales")
library(scales)

g <- crosstalk::SharedData$new(gapminder, ~continent) #Creates slider widget
 
gg <- ggplot(g, aes(gdpPercap, lifeExp, color = continent, frame = year)) +
  #gdpPercap is on the x-axis and lifeExp is on the y-axis; color represents continent
  geom_point(aes(size = pop, ids = country)) + 
  #size of bubbles represents population by country
  #geom_smooth(se = FALSE, method = "lm") + #this is an optional smoothing line
  labs(x = "Income",
    y = "Life Expectancy",
    color = "grey",
    title = "Income Level 1 <> Level 2 <> Level 3 <> Level 4") +
  scale_y_continuous( breaks = c(10, 20, 30, 40, 50, 60, 70, 80, 90) ) + 
  scale_x_log10( breaks = c(500, 1000, 2000, 4000, 8000, 16000, 32,000, 64000, 128000) ) +
  theme_classic() +
  scale_fill_manual( breaks = c("Asia","Africa","Americas","Europe","Oceania") , 
                      values = c("pink","cyan","chartreuse1","gold","firebrick1") ) +
  theme(plot.background = element_blank() ) +
  annotate("text", x = 5000, y = 50, label = "YEARS 1952 - 2007",
             hjust=15, vjust=15, col="black",
             fontface = "bold", alpha = 0.5)

ggplotly(gg) %>% 
   highlight("plotly_hover")

```
Author: Daniel Posmik (posmikdc@mail.uc.edu)